---
title: "Assignment 5: Data import (readr) and tidy data (tidyr)"
Author: Shitaye
Date: 02/19/2020
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**Instructions: Please read through this before you begin**

*This homework is due by **10pm on Monday 10/26/20**

*Please **reproduce this markdown template**. Pay attention to all the formating in this file, including bullet points, bolded characterrs, inserted code chunks, headings, text colors, blank lines, etc.

*For the first two excercises in this assignment, **import** data files into R and **parse** them properly as instructed. **Tidy and/or **transform** the data frames when appropriate.

  + We have not extensively covered parsing data in lecture, but you can find more information about here.
  
  + Briefly, you can choose to use the `col_types` argument in a`read_*()` function to parse the data during    import.
  + In many cases, you can also choose to parse the data using one of the `parse_*()` functions nested within    a `mutate()` after importing the data.
  
  + In many cases, you can choose to parse the data using one of the `parse_*()` functions nested within a      `mutate()` function after importing the data.
  
* For the next exercise, you will explore a dataset with details about passengers on the Titanic. First, **answer the questions below and use figures or tables to support your answer** Then, **explore the dataset on your own** using the data transformation and visualization skills that you have learned in this so far. 

    + For this excercise, pleas make sure to put some thought into **editing the aesthetics of your figures and tables** to make them easier to understand and nicer to look at (e.g. choose the most appropriate geometric object, aesthetic mapping, facetting, postion adjustment; add meaningful axis lables, figure titles legend titile; change the background; be creative; and etc.)
    
    
* Please note that the **excercise 4 is optional**. It is a good exercise for practicing your data tidying and exploration skills through, so we encourage you to give it a try if time allows.

* Please **reproduce the tables and plots exactly as shown in this markdown file**.

* When a verbal response is needed, answer by editing the part in the R markdown template where it says "Write your response here" .

* Have all your code embedded within the R markdown file, and show both of your **code** and **plots** in the    kitted markdown file.
* Use R Markdown functionalities to **hide message and warnings when needed**. (Suggestion: messages and     warnings can often be informative and important, so please examine them carefylly and only turn them off when you finish the exercise).

* Please name your R markdown file `assignment_5.Rmd` and the knitted markdwon file `assignment_5.md`. Please upload both files using your personal GitHub repository for this class.

* To start, first load all the required packages with the following code. Install them if they are not installed yet.

```{r, message = FALSE}

library(tidyverse)
library(knitr)
```



##**Exercise 1. Tibble and Data Import

Import data frames listed below into R and parse the columns appropraitly when need ed. Watch out for the formating oddities of each dataset. Print the results with `kable()`.

1.1 **Create the following tibble manually, first using `tribble()` and then using `tibble()`. Print both results.**

```{r}
tribble(~a,~b, ~c,
        1, 2.1, "apple",
        2, 3.2, "orrange"
        ) %>% 
  kable()

```

    
  
```{r}
tibble(a = c(1,2),
       b = c(2.1, 3.2), 
       c = c("apple", "orrange"))%>% 
  kable()
```
  
  
**1.2 import `https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset2.txt` into R. Change the column names into “Name”, “Weight”, “Price”.**
  
```{r}

data_Fruit<-read.table("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset2.txt", sep = ",") 

colnames(data_Fruit) <- c("Name", "Weight", "Price")
kable(data_Fruit)

```
  

**1.3 Import `https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset3.txt` into R . Watch out for the first few lines, missing values, separators, quotation marks, and deliminaters.**

```{r}
read.table("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset3.txt", sep, ",")
```




  
  
  
