---
title: "Assignment 5: Data import (readr) and tidy data (tidyr)"
Author: Shitaye
Date: 02/19/2020
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#**Instructions: Please read through this before you begin**

* This homework is due by **10pm on Monday 10/26/20**

* Please **reproduce this markdown template**. Pay attention to all the formating in this file, including bullet points, bolded characterrs, inserted code chunks, headings, text colors, blank lines, etc.

* For the first two excercises in this assignment, **import** data files into R and **parse** them properly as instructed. **Tidy and/or **transform** the data frames when appropriate.

  + We have not extensively covered parsing data in lecture, but you can find more information about here.
  
  + Briefly, you can choose to use the `col_types` argument in a`read_*()` function to parse the data during    import.
  + In many cases, you can also choose to parse the data using one of the `parse_*()` functions nested within    a `mutate()` after importing the data.
  
  + In many cases, you can choose to parse the data using one of the `parse_*()` functions nested within a      `mutate()` function after importing the data.
  
* For the next exercise, you will explore a dataset with details about passengers on the Titanic. First, **answer the questions below and use figures or tables to support your answer** Then, **explore the dataset on your own** using the data transformation and visualization skills that you have learned in this so far. 

    + For this excercise, pleas make sure to put some thought into **editing the aesthetics of your figures and tables** to make them easier to understand and nicer to look at (e.g. choose the most appropriate geometric object, aesthetic mapping, facetting, postion adjustment; add meaningful axis lables, figure titles legend titile; change the background; be creative; and etc.)
    
    
* Please note that the **excercise 4 is optional**. It is a good exercise for practicing your data tidying and exploration skills through, so we encourage you to give it a try if time allows.

* Please **reproduce the tables and plots exactly as shown in this markdown file**.

* When a verbal response is needed, answer by editing the part in the R markdown template where it says "Write your response here" .

* Have all your code embedded within the R markdown file, and show both of your **code** and **plots** in the    kitted markdown file.
*Use R Markdown functionalities to **hide message and warnings when needed**. (Suggestion: messages and     warnings can often be informative and important, so please examine them carefylly and only turn them off when you finish the exercise).

* Please name your R markdown file `assignment_5.Rmd` and the knitted markdwon file `assignment_5.md`. Please upload both files using your personal GitHub repository for this class.

* To start, first load all the required packages with the following code. Install them if they are not installed yet.

```{r, message = FALSE}

library(tidyverse)
library(knitr)
```



##**Exercise 1. Tibble and Data Import

Import data frames listed below into R and parse the columns appropraitly when need ed. Watch out for the formating oddities of each dataset. Print the results with `kable()`.

1.1 **Create the following tibble manually, first using `tribble()` and then using `tibble()`. Print both results.**

```{r}
tribble(~a,~b, ~c,
        1, 2.1, "apple",
        2, 3.2, "orrange"
        ) %>% 
  kable()

```

    
  
```{r}
tibble(a = c(1,2),
       b = c(2.1, 3.2), 
       c = c("apple", "orrange"))%>% 
  kable()
```
  
  
**1.2 import `https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset2.txt` into R. Change the column names into “Name”, “Weight”, “Price”.**
  
```{r}

data_Fruit<-read.table("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset2.txt", sep = ",") 

colnames(data_Fruit) <- c("Name", "Weight", "Price")
kable(data_Fruit)

```
  

**1.3 Import `https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset3.txt` into R . Watch out for the first few lines, missing values, separators, quotation marks, and deliminaters.**

```{r}

read_table("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset3.txt")#, row.names = 1)# sep = ",")
Fruit<-read_csv("Name, Weight,Price
         apple,1,2.9
         orange, 2, NA
         durian, NA, 19.9")
kable(Fruit)

```



**1.4 Import `https://raw.githubsercontent.com/nt246/NTRES6940-science/master/datasets/dataset4.txt` into R. Watch out for comments, units, and decimal markes(which are `,` in this case).**


```{r}
#read_csv("https://raw.githubsercontent.com/nt246/NTRES6940-science/master/datasets/dataset4.txt")#, sep =",")
#hacky hour not working
```

**1.5 import `https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset5.txt` into R. Parse the columns properly. Write this imported and parsed data frame into a new csv file named `dataset5_new.csv` in your `problem_sets` folder.**

```{r}

Expdate<-read.table("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/dataset5.txt")
colnames(Expdate) <- Expdate[1,]
Expdate<-Expdate[-1,]
kable(Expdate)
write.csv(Expdate, file = "dataset5_new.csv") #imported dataframe


```

It is not clear how to organize the dataframe after parsing

#**Exercise 2.Weather station**

This dataset contains the weather and air quality data collected by a weather station in Taiwan. It was obtained from the Environmental Protection Administration, Excutive Yuan, R.O.C. (Taiwan).

**2.1 The text file `https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/2015y_Weather_Station_notes.txt` contains desciptions of different variables collected by the station. Import it into R and print it in a table as shown below.**



```{r}

weather<-read_delim("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/2015y_Weather_Station_notes.txt",
       delim = "-" )

kable(weather)

```


`#` indicates invalid value by equipment inspection

`*` indicates invalid value by program inspection

`x` indicates invalid value by human inspection

`NR` indicates no rainfall

blank indicates no data


**2.2 Import `https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/2015y_Weather_Station.csv` in to R. As you can see, this dataset is a classic example of untidy data: values of a variable (i.e. hour of the day) are stored as column names; variable names are stored in the `item' column. Clean this dataset up by turning it into a tidy format. Also, parse the `date` variable into date format and parse `hour` into time. Turn all invalid values into `NA` and turn`NR` in rainfall into `0`. Parse all values into numbers. Show the first 6 rows and 15 columns of this cleaned dataset, as shown below. (Hint: you don't have to do these tasks in the given order.)**

Before cleaning:

```{r}

rainfall<-read_csv("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/2015y_Weather_Station.csv") 

kable((rainfall)[1:6,1:15])



```

After cleaning


```{r, message =FALSE, warning=FALSE}
weather_new <- pivot_longer(rainfall, cols = 4:27, names_to = "hour", values_to = "value") %>%
  mutate(value=ifelse(value=="NR", 0, value), value=parse_double(value),
         hour=parse_time(hour, "%H")) %>%
  pivot_wider(names_from = item, values_from = value)
kable((weather_new)[1:6,1:15])

```


**2.3 Using this cleanded dataset, plot the daily variation in ambient temperature on september 25, 2015, as shown below.**

```{r}
sept<-weather_new %>% 
  filter(date=="2015-09-25")
ggplot(sept, mapping = aes(x = hour, y = AMB_TEMP))+
  geom_line()

```



**2.4 Plot the daily average ambient temperature throught the year with a continuous line, as shown below.**


```{r, message=F}

weather_new %>% 
  group_by(date) %>% 
  summarise(AMB_TEMP = mean(AMB_TEMP, na.rm = T)) %>% 
  ggplot(mapping = aes(x=date, y= AMB_TEMP))+
geom_line()
  
```




**2.5 Plot the total rainfall per month in a bar chart, as shown below.(Hint: separating date into three columns might be helpful.)**

```{r, message=FALSE}

library(lubridate)
```



```{r,message=F}

weather_new %>% 
  mutate_at(vars(date),funs(year,month, day))%>% 
  group_by(month)%>% 
  summarize(MonthlyRainfall= sum(RAINFALL,na.rm = T))%>% 
  ggplot(mapping = aes(x=month, y = MonthlyRainfall))+
  geom_bar(stat = "identity")

```


**2.6 Plot the per hour variation in PM2.5 in the first week of September with a continuous line, as shown below. (Hint: uniting the date and hour and parsing the new variable might be helpful.)**

```{r}

year_month_day<-weather_new %>% 
  mutate_at(vars(date),funs(year,month, day))


wk1sept<- year_month_day %>% 
  filter(month==9 & day ==1:8) 
  #mutate(time=str_c(month,day),time = md(time) ) # "%b/%d"))
  
```


#**Excercise 3: Titanic passengers**

**This dataset contains information about a subset of the passengers who were abroad the Titanic when it sank in 1912.**

```{r, message = F}
titanic<-read_csv("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/Titanic.csv")
kable(head(titanic))
```


And here are descriptions of the variables in the dataset

```{r, message= F}

notes<-read_csv("https://raw.githubusercontent.com/nt246/NTRES6940-data-science/master/datasets/Notes.csv")
kable(notes)

```

##**Part 1.**

**Answer the questions below** and use **figures or tables** to support your answer. Choose the most appropriate type of figure or table for each question.

**3.1 According to Wikipedia, there was an estimated 2,224 passengers and crew onboard the Titanic when it sank. How many of them do we have information on in this dataset? How many of them survived and how many did not? What is the overall survival rate?**

```{r}

titanic %>% 
  summarize(total_individuals=sum(Survived==1|Survived==0),
            total_survived =sum(Survived==1),total_notsurvived=sum(Survived ==0), 
            survival_rate= total_survived/total_individuals) %>% 
kable()

```


We have the information of 891 passengers. 342 individuals survived and 549 notsurvived and the survivale rate was 0.39.


**3.2 How many passengers on the Titanic were males and how many were females? What do you find when you break it down by ticket class?**

```{r}

titanic %>% 
  summarize(total_male=sum(Sex=="male"), total_female =sum(Sex=="female")) %>% 
  kable()

titanic %>% 
  summarize(ticketclass1=sum(Pclass==1),
            ticketclass2=sum(Pclass==2),ticketclass3=sum(Pclass==3)) %>% 
  kable()


```

314 were female and 577 were male

**3.3 How many passengers of each sex survived and how many of them did not? What is the survival rate for passengers of each sex?**

```{r}

 titanic %>% 
  filter(Sex=="male") %>% 
  summarise(Male_survived = sum(Survived==1), Male_notsurvived= sum(Survived==0),survival_rateMale = Male_survived/sum(Male_survived, Male_notsurvived)) %>% 
  kable()
 titanic %>% 
  filter(Sex=="female") %>% 
  summarise(Female_survived = sum(Survived==1), Female_notsurvived=sum(Survived==0), 
            survival_ratFemale= Female_survived/sum(Female_survived,Female_notsurvived)) %>% 
    kable()




```


Write your response here

**3.4 How many passengers do we have age information for (including estimated age)? For how many is the age information missing? What is the age distribution for passengers whose age information is available?**

```{r}
## Write your code here
```


Write your response here

**3.5 Show the age distribution per ticket class, per sex. What do you find?**

```{r}
## Write your code here
```

Write your response here

**3.6 What is the age distribution of passengers who survived vs. those who did not? What hypothesis can you come up with when comparing these two distributions?**


```{r}
## Write your code here
```


Write your response here


**3.7 In this dataset, the Fare variable does not represent the fare per person. Instead, each ticket number has a corresponding fare, and some passengers share one single ticket number. Therefore, the Fare variable is the total fare for a group of passengers sharing the same ticket number. Knowing this, calculate the average fare per person. (You don’t need to show a table or a figure for this question, just show the code for the calculation)**

```{r}
## Write your code here
```

Write your response here

**3.8 Show the distribution of the number of family members (including siblings, spouses, parents, and children) that each passenger was accompanied by. Were most passengers travelling solo or with family?**


```{r}
## Write your code here
```

Write your response here

**3.9 Which ticket class did most of the largest families get? And which ticket class has the lowest proportion of female passengers who travelled solo out of all the female passengers in that class?**

```{r}
## Write your code here
```

Write your response here

**3.10 Do the port of embarkation matter for the survival rates of passengers? Why might this be the case?**

```{r}
## Write your code here
```

Write your response here









  
